---
title: "HPGC Fourball Match Play 2016"
output: html_document
---

```{r echo=FALSE}
library(knitr)
```



Last updated: `r format(Sys.time(),"%b %d, %Y %H:%M")`

##Table of Contents  

1.  [Pool Play](#pp)  
2.  [Tournament Format](#format)  
3.  [Roster (including current handicaps)](#roster)  
  

##Pool Play  
<a id="pp"></a>  

###Recent Results  

*  7/3 - Schneider/Drish def. Beatty/Schroeder, 10 - 8 (3&2)  
*  7/5 - Ota/Yeh halve with Dhadal/Tao, 9 - 9 (A/S)
*  7/9 - Lee/Kahn def. Brennan/Henderson, 11.5 - 6.5 (5&3)  
*  7/17 - Kalamaras/Kretz def. Beatty/Schroeder, 10 - 8 (3&2)  
*  7/17 - Simpson/Solomon def. Henneberry/Ashburner, 11 - 7 (4&2)  


###Current Standings  

```{r echo=FALSE}
pooler <- read.csv("poolPlay.csv")
names(pooler) = c("Team","Pool","Player 1","Player 2","Wins","Losses","Ties","Points","Holes Won")

## Schneider/Drish defeat Beaty/Schroeder, 10-8
pooler[1,"Wins"] <- pooler[1,"Wins"] + 1
pooler[1,"Points"] <- pooler[1,"Points"] + 3
pooler[1,"Holes Won"] <- pooler[1,"Holes Won"] + 10
pooler[7,"Losses"] <- pooler[7,"Losses"] + 1
pooler[7,"Holes Won"] <- pooler[7,"Holes Won"] + 8

## Lee/Kahn def. Brennan/Henderson, 11.5 - 6.5
pooler[3,"Wins"] <- pooler[3,"Wins"] + 1
pooler[3,"Points"] <- pooler[3,"Points"] + 3
pooler[3,"Holes Won"] <- pooler[3,"Holes Won"] + 11.5
pooler[18,"Losses"] <- pooler[18,"Losses"] + 1
pooler[18,"Holes Won"] <- pooler[18,"Holes Won"] + 6.5

## Kalamaras/Kretz defeat Beaty/Schroeder, 10-8
pooler[17,"Wins"] <- pooler[17,"Wins"] + 1
pooler[17,"Points"] <- pooler[17,"Points"] + 3
pooler[17,"Holes Won"] <- pooler[17,"Holes Won"] + 10
pooler[7,"Losses"] <- pooler[7,"Losses"] + 1
pooler[7,"Holes Won"] <- pooler[7,"Holes Won"] + 8

## Ota/Yeh halve with Dhadal/Tao, 9 - 9
pooler[12,"Ties"] <- pooler[12,"Ties"] + 1
pooler[12,"Points"] <- pooler[12,"Points"] + 1
pooler[12,"Holes Won"] <- pooler[12,"Holes Won"] + 9
pooler[14,"Ties"] <- pooler[14,"Ties"] + 1
pooler[14,"Points"] <- pooler[14,"Points"] + 1
pooler[14,"Holes Won"] <- pooler[14,"Holes Won"] + 9 

## Simpson/Solomon def. Henneberry/Ashburner, 11 - 7
pooler[11,"Wins"] <- pooler[11,"Wins"] + 1
pooler[11,"Points"] <- pooler[11,"Points"] + 3
pooler[11,"Holes Won"] <- pooler[11,"Holes Won"] + 11
pooler[20,"Losses"] <- pooler[20,"Losses"] + 1
pooler[20,"Holes Won"] <- pooler[20,"Holes Won"] + 7  

```


###Group A  
```{r echo=FALSE, results="asis"}

kable(pooler[pooler$Pool=="A",-2],row.names=FALSE)
```  


###Group B  
```{r echo=FALSE, results="asis"}

kable(pooler[pooler$Pool=="B",-2],row.names=FALSE)
``` 


###Group C  
```{r echo=FALSE, results="asis"}

kable(pooler[pooler$Pool=="C",-2],row.names=FALSE)
``` 

###Group D  
```{r echo=FALSE, results="asis"}

kable(pooler[pooler$Pool=="D",-2],row.names=FALSE)
``` 

###Group E  
```{r echo=FALSE, results="asis"}

kable(pooler[pooler$Pool=="E",-2],row.names=FALSE)
``` 

##Tournament Format, Rules, and Schedule  
<a id="format"></a>  

###Match Format  
This tournament is standard fourball match play--teams of two play against eachother in match play.  Both team members play their own ball, with the lower (net) score of the partners being the team's score for that hole.  If teams tie on a hole the hole is halved, there are no carry-overs  

###Location  
All matches will be played at TPC Harding Park  

###Tees  
All matches will be played from the white tees  

###Handicaps  
All matches will be played net.  We will follow the USGA method for assigning handicaps for net fourball play--the lowest course handicap in the foursome will play off scratch, with each of the remaining players playing to the difference between their course handicap and the lowest.  For example, if the four players have course handicaps of 4, 7, 7, and 8, the 4 handicap will get 0 strokes, the two 7's will get strokes on handicap holes 1 - 3, and the 8 handicap will get strokes on handical holes 1 - 4.  

The handicaps in play for each match will be the course handicaps based on full GHIN indexes as of the date of the match.  For transparency, indices for all players will be updated bi-weekly with the GHIN updates on this site in the roster section [here](#roster)

###Rules  
USGA rules of golf and the [HPGC local tournament rules](http://www.golffusion.com/hardingparkgolf/content/1576305/Local-Tournament-Rules) will govern all play

###Schedule  
Pool play will take place from June through the end of September, with playoffs to begin in October.  All matches, including the playoffs, will be self-scheduled, and players are reponsible for their own greens fees.  Early weekend times are always difficult at TPC Harding Park, so I recommend booking early as cancellation is free.

###Pool Play  
Teams have been grouped into five pools of four.  The first part of the tournament is round robin pool play--each team will play each otehr team in their pool once.  

During pool play, matches will continue for all 18 holes, regardless of who wins.  Each hole won will be worth one point for the winning team, while halved holes will be worth a half point for both teams.  The team with the most holes won wins regardless, but total holes won will be important for breaking ties in the likely event that teams have the same record 

###Order of finish  
Final standings within the pool will be based on:  

1.  Win/Loss record:  Wins are worth 3 points, ties are worth 1  
2.  If tied on #1, head-to-head record
2.  If tied on #2, total holes won  

###Playoffs  
Eight teams will make the playoffs, the five pool winners plus three wildcard teams based on (1) pool play match record or (2) total holes won during pool play.

Playoffs will be single elimination play.  I will work with management at TPC Harding Park to see if we can use sudden-death extra holes for these matches in the case of a half, if not we will use a tie-breaking system that is TBD.

##Roster  
<a id="roster"></a>
```{r echo=FALSE}

ch <- function(idx) {
    isNH <- idx=="NH"
    isPlus <- (substr(idx,1,1)=="+")
    rating <- 124
    hin <- numeric(length(idx))
    hin[!isNH] <- as.numeric(idx[!isNH])
    chi <- character(length(idx))
    chi[!isNH] <- as.character(round(rating/113*hin[!isNH],0))
    chi[isPlus&&chi!="0"] <- paste("+",chi[isPlus])
    chi[isNH] <- "0"
    chi
    
}

roster <- read.csv("../rosterPools.csv")
index <- read.csv("../indexResults.csv", colClasses = c("numeric","character","character"))
roster <- merge(roster, index, by.x="ghin",by.y="ghin")
roster$index[roster$ghin==9999999] <- "NH"

roster <- roster[,c("Name","teamNo","pools","index","asOf")]

roster$'Course Handicap' <- ch(roster$index)
names(roster) <- c("Player","Team","Pool","Handicap Index","Current As Of", "Course Handicap")
roster <- roster[order(roster$Team),]
kable(roster,row.names=FALSE)
```
